Scalar on Function Regression

### 2022-01-06 Presentation/slides.tex
Notation

		\begin{itemize}
			\item Near-infrared (NIR) spectroscopy enables fast diagnostics by using the NIR region of the electromagnetic spectrum
			\item Suited for field-monitoring / on-line analysis and diagnostics of e.g: prediction of octane ratings!
			\item Spectroscopy results in high-dimensional dataset.

			\item
			This set of measurements along a continuum can be viewed as set of smooth spectral curves

			\item
			Regression to determine relationship between octane rating and spectral curves
			\end{itemize}
	\end{frame}

	\begin{frame}{Theory}

		\begin{itemize}
			\item
			A simple functional dataset is given by $x_{n}(t_{j,n}) \in \mathbb{R}, t_{j,n} \in [T_1, T_2], n = 1,2,...,N, j = 1,2,..., J_n  $
			\item Continuous underlying process, where $x_n(t)$ exists at any point $t$, but is only observed at $x_{n}(t_{j,n})$, eg.: growth curves, financial data, human perception (music pitch,...).

			\item To abstract information from the curves, they must be interpretable!

			\end{itemize}
	\end{frame}

@@

	\begin{frame}{Estimation via Basis Representation}
		Assume the following data generating process
		$$Y(\omega) = \alpha + \int_{0}^{1} \beta(s) F(\omega)(s) \mathrm{d}s + \epsilon(\omega)$$
		\begin{itemize}
			\item $Y$ and $\epsilon$ realize in $\mathbb{R}$ and $F$ realizes in $\mathbb{L}^2[0,1]$
		\end{itemize}
		\vspace{0.2cm}

		Assume that we have a data set containing observations each of which is made up of:
		\begin{itemize}
			\item $y_i$: a scalar realization of $Y$
			\item $f_i(t)$: a realization of $F$
		\end{itemize}

@@

	\begin{frame}{Estimation via Basis Representation}
		$$Z_j(\omega) = \int_{0}^{1} F(\omega)(s) b_j(s)\mathrm{d}s$$ 
		This is a scalar random variable leading to the following transformation
		$$Y(\omega) = \alpha + \sum_{j = 1}^{\infty} \psi_j Z_j(\omega) + \epsilon(\omega)$$
		Each combination of observation $f_i(t)$ and deterministic basis function $b_j(t)$ effectively gives us a realization of this random variable.
		$$Z_{i,j} = \int_{0}^{1} f_i(s) b_j(s)\mathrm{d}s$$

@@

		Coefficients can then be estimated using theory from multivariate regression leading to an estimated coefficient vector $\hat{\psi}_L \in \mathbb{R}^L$.
	\end{frame}

@@

	\begin{frame}{Karhunen-Lo\'{e}ve Expansion}
		A random function $F$ can be expressed in terms of its mean function and its Eigenfunctions:
		$$F(\omega)(t) = \mu(t) + \sum_{j = 1}^{\infty} \xi_j(\omega) \nu_j(t)$$

@@

	\begin{frame}{FPCA - Plots}
		\begin{minipage}[t]{0.79\textwidth}
			\includegraphics[width = \textwidth]{../Graphics/principal_components.pdf}
			\includegraphics[width = \textwidth]{../Graphics/pc_approx.pdf}
		\end{minipage}
		\begin{minipage}[t]{0.19\textwidth}
			Lorem Ipsum Dolor Sit Amet
		\end{minipage}
	\end{frame}

	\begin{frame}{Simulation Setup \& Application}
		\begin{itemize}
		\item
		 Use the gasoline dataset (NIR-spectroscopy, 60 $\times$ 401) to predict octane ratings.
		\item 
		Generate new curves from gasoline dataset, motivated by Karhunen-Loeve expansion:

		$$F^{new}(\omega)(t) = \mu^{NIR}(t) + \sum_{j = 1}^{\infty} \xi_j^{new} \nu_j(t)^{new}$$ \\
		where \\
		 \vspace{0.1cm}
		$$\xi_{j}^{new} \in  \mathcal{N}_{n}(0,\Sigma)$$ \\
		and $\nu_j(t)^{new}$ be a $grid \times n_{FPC}$ matrix.

		\end{itemize}
	\end{frame}

	\begin{frame}{Simulation Setup \& Application cont.}
		\begin{itemize}
		\item
		Following Reiss and Ogden (2007), let $f_1$ and $f_2$ be two true coefficient functions that differ in smoothness: 
		\vspace{0.4cm}

   $f_1 = 2\sin(0.5\pi t) + 4\sin(1.5 \pi t) + 5\sin(2.5 \pi t)$\\
   \vspace{0.1cm}

    $	f_2 = 1.5^{\frac{-0,5(t-0.3)^2}{0.02^2}} - 4^{\frac{-0,5(t-0.45)^2}{0.015^2}} +  8^{\frac{-0,5(t-0.6)^2}{0.02^2}} -  1^{\frac{-0,5(t-0.8)^2}{0.03^2}}$

		\end{itemize}
		\vspace{0.1cm}
		\begin{figure}
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[height=3cm]{smooth_function.png}
  \captionof{figure}{$f_1$, smooth function}
  \label{fig:test1}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[height=3cm]{bumpy_function.png}
  \captionof{figure}{$f_2$, bumpy function}
  \label{fig:test2}
\end{minipage}
\end{figure}

@@

	\begin{frame}{Simulation Setup \& Application cont.}
		\begin{itemize}
		\item
    		Let \\
    		 \vspace{0.1cm}

    		$Y_{1,f} = \mathbf{NIR \times f}  + \mathcal{N}(0,1) \times var(NIR \times f)/0.9 - var(NIR \times f)$, \\
    		  $Y_{2,f} = \mathbf{NIR \times f} + \mathcal{N}(0,1) \times var(NIR \times f)/0.9 - var(NIR \times f)$ be two responses for $f \in f_1, f_2$.		
    		 	\item
    		 	 Four combinations with different number of cubic basis-function $n_{basis} \in (5,6,...,25)$ to perform regression using basis expansion and the FPCR approach.
			\item Compare results via criteria (CV, Mallows CP,...)
			\item 
			{\color{green} add results here!}

		\end{itemize}
	\end{frame}

>>>>

		\begin{itemize}
			\item Near-infrared (NIR) spectroscopy enables fast diagnostics by using the NIR region of the electromagnetic spectrum
			\item Suited for field-monitoring / on-line analysis and diagnostics of e.g: prediction of octane ratings!
			\item Spectroscopy results in high-dimensional dataset.	
			\item This set of measurements along a continuum can be viewed as set of smooth spectral curves
			\item Regression to determine relationship between octane rating and spectral curves
			\end{itemize}
	\end{frame}

	\begin{frame}{Theory}
		A simple functional dataset is given by 
		$$\{x_{i}(t_{j,i}) \in \mathbb{R} \: \vert \:  t_{j,n} \in [T_1, T_2], \; i = 1,2,...,N, \; j = 1,2,..., J_i  \}$$

		\begin{itemize}
			\item Continuous underlying process, where $x_n(t)$ exists $\forall t \in [T_1, T_2]$
			\item Only observed at $x_{n}(t_{j,n})$
			\item Growth curves, financial data, human perception (pitch), ...
			\item To abstract information from the curves, they must be interpretable!
			\end{itemize}
	\end{frame}

@@

	\begin{frame}{Estimation via Basis Representation}
		Assume the following \textbf{data generating process}
		$$Y(\omega) = \alpha + \int_{0}^{1} \beta(s) F(\omega)(s) \mathrm{d}s + \epsilon(\omega)$$
		\begin{itemize}
			\item $Y(\omega)$ and $\epsilon(\omega)$ realize in $\mathbb{R}$ and $F(\omega)$ realizes in $\mathbb{L}^2[0,1]$
		\end{itemize}
		\vspace{0.2cm}

		Assume that we have a data set containing observations each of which is made up of:
		\begin{itemize}
			\item $y_i$: a scalar realization of $Y(\omega)$
			\item $f_i(t)$: a realization of $F(\omega)$
		\end{itemize}

@@

	\begin{frame}{Estimation via Basis Representation}
		$$Z_j(\omega) = \int_{0}^{1} F(\omega)(s) b_j(s)\mathrm{d}s$$ 
		This is a \textbf{scalar random variable} leading to the following transformation
		$$Y(\omega) = \alpha + \sum_{j = 1}^{\infty} \psi_j Z_j(\omega) + \epsilon(\omega)$$
		Each combination of observation $f_i(t)$ and deterministic basis function $b_j(t)$ effectively gives us a realization of this random variable.
		$$Z_{i,j} = \int_{0}^{1} f_i(s) b_j(s)\mathrm{d}s$$

@@

		Coefficients can then be estimated using theory from \textbf{multivariate regression} leading to an estimated coefficient vector $\hat{\psi}_L \in \mathbb{R}^L$.
	\end{frame}

@@

	\begin{frame}{Karhunen-Lo\'{e}ve Expansion}
		A random function $F(\omega)$ can be expressed in terms of its mean function and its Eigenfunctions:
		$$F(\omega)(t) = \mu(t) + \sum_{j = 1}^{\infty} \xi_j(\omega) \nu_j(t)$$

@@

	\begin{frame}{FPCA - Plots}
		\begin{minipage}[c]{0.78\textwidth}
			\includegraphics[width = \textwidth]{../Graphics/principal_components.pdf}
			\includegraphics[width = \textwidth]{../Graphics/pc_approx.pdf}
		\end{minipage}
		\begin{minipage}[c]{0.19\textwidth}
			The math is for intuition. In practice there are problems and the fpc's are derived differently.
		\end{minipage}
	\end{frame}

	\begin{frame}{Simulation Setup \& Application}
		\begin{itemize}
			\item Use the \textbf{gasoline dataset} (NIR-spectroscopy, 60 $\times$ 401) to predict octane ratings.
			\item Generate \textbf{similar curves} from gasoline dataset:
		\end{itemize}

		$$\tilde{F}(\omega)(t) = \hat{\mu}(t) + \sum_{j = 1}^{\infty} \tilde{\xi}_j(\omega) \hat{\nu}_j(t)$$

		\begin{itemize}
			\item $\tilde{\xi}_{j} \sim \mathcal{N}(0,\hat{\lambda}_j)$ (simplification - the $\xi_{j}$ do not follow a normal)
			\item $\tilde{F}(\omega)(t)$, $\hat{\mu}(t)$ and $\hat{\nu}_j(t)$ are approximated as vectors in $\mathbb{R}^{401}$.
		\end{itemize}

@@

	\begin{frame}{Simulation Setup \& Application cont.}
		%\begin{itemize}
		%\item
		Following \textbf{Reiss and Ogden (2007)}, let $f_1(t)$ and $f_2(t)$ be two coefficient functions: 
		%\vspace{0.2cm}

   		%$f_1(t) = 2\sin(0.5\pi t) + 4\sin(1.5 \pi t) + 5\sin(2.5 \pi t)$\\
   		%\vspace{0.1cm}

        %$f_2(t) = 1.5 \exp(\frac{-0.5(t-0.3)^2}{0.02^2}) - 4 \exp(\frac{-0.5(t-0.45)^2}{0.015^2}) +  8 \exp(\frac{-0.5(t-0.6)^2}{0.02^2}) - \exp(\frac{-0.5(t-0.8)^2}{0.03^2})$

		%\end{itemize}
		\vspace{0.1cm}
		\begin{figure}
			\centering
			\begin{minipage}{.5\textwidth}
				\centering
  				\includegraphics[width=\textwidth]{smooth_function.png}
  				\caption{$f_1(t)$, smooth function}
  				\label{fig:test1}
			\end{minipage}%
			\begin{minipage}{.5\textwidth}
	  			\centering
  				\includegraphics[width=\textwidth]{bumpy_function.png}
  				\caption{$f_2(t)$, bumpy function}
  				\label{fig:test2}
			\end{minipage}
		\end{figure}

	\begin{frame}{Simulation Setup \& Application cont.}
		Let 
		$$Y_{1,f} = \langle NIR, f\rangle + Z  \frac{var(\langle NIR, f\rangle)}{0.9} - var(\langle NIR, f\rangle)$$ 
		$$Y_{2,f} = \langle NIR, f\rangle + Z  \frac{var(\langle NIR, f\rangle)}{0.9} - var(\langle NIR, f\rangle)$$

		where $Z \sim \mathcal{N}(0,1)$ be two responses for $f \in \{f_1(t), f_2(t)\}$.	
		\begin{itemize}
    		\item Four combinations with different number of cubic basis-function $n_{basis} \in (5,6,...,25)$ to perform regression using basis expansion and the FPCR approach.
			\item Compare results via criteria (CV, Mallows CP,...)
			\item {\color{green} add results here!}
		\end{itemize}
	\end{frame}
